Determining the type of every positional distinct array is done by first determining the key type, $t\in T$, then determining the type of the values, and from this deduce the type as either List, Map or Object as defined in the beginning of this chapter. Since an array can change type during a program, the type of the keys is more accurately a set of types. If an array is observed with different key types throughout the analysis, then the key type of the array is the set of these types. E.g, let an array be observed at one point with type List and later with type Sparse List, then the type of the array is $\{$List, Sparse List$\}$.

Detecting the key type for each array is done by traversing the file from top to bottom inspecting each line. If a line contains an array $a$ the type of the line is associated with the corresponding ID of the array.

Detecting the type of values is also done by traversing through the log file from top to bottom. Here the reads and writes from and to the arrays are used to determine the types of the values. This is done by associating all the types of the values read/written with the respective array. 

For every array with key and value type information, it is now possible to determine whether it is a list, map, object or uncategorizable. Given an array, $a$, with type information, if $a$ has multiple key types, it is considered uncategorizable. Else if $a$ contains values of a single type, it is either a map or a list, depending on the key type. If the values have multiple types, then $a$ is an object.

When the types of the arrays are determined, we can analyse the operations used with each type of array. This is done by once again iterating through the log file and associating line types with the type of the arrays.