\section{Control-flow graph}
\label{sec:cfg}
\tikzstyle{subtree} = [rectangle, draw, fill=gray!20,  minimum width=4em, text badly centered, minimum height=3em]
\tikzstyle{node} = [draw, circle, minimum height=2em]
\tikzstyle{emptynode} = [draw, circle]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{entry} = [draw=none,fill=none]
\tikzstyle{exit} = [draw=none,fill=none]
\tikzstyle{dots} = [draw=none,fill=none]
\newcommand{\subt}[1]{\llbracket #1\rrbracket}

The control-flow graph defines the basic instructions of a program and how data flows between those instructions. Nodes correspond to instructions and edges to data flow. The graph is built recursively by defining sub graphs corresponding to grammar \ref{gramm:p0}. Table \ref{tab:nodes} gives a description of sub-graphs and their respective arguments.

\bgroup
\def\arraystretch{1.5}
\begin{table}
\centering
\begin{tabularx}{\textwidth}{c|X}
 Symbol & Description \\\hline\hline
 $t \in \text{Temps}$ & A temporary variable name used to pass evaluated values between CFG nodes \\
 $h \in \text{HeapTemps}$ & A temporary heap variable name used to pass a set of heap locations between CFG nodes \\
 $c \in \text{Temps}\bigcup\text{HeapTemps}$ & Either of the above two types of variable names \\
 $\subt{E}(t)$ & Sub graph for expression $E$ with evaluation result stored in $t$ \\
 $\subt{R}(h)$ & Sub graph for reference expression $R$ with evaluated heap-location set stored in $h$ \\
 $\subt{V}(h)$ & Sub graph for a variable expression $V$ with evaluated heap-location set stored in $h$ \\
 $\subt{T}(c)$ & Sub graph for either an expression, a reference expression or a variable expression with $c$ being the corresponding $t$ or $h$ as described above \\
 $\subt{S}$ & Sub graph for statement $S$ 
\end{tabularx}
\caption{Symbols used in the control-flow graph}
\label{tab:nodes}
\end{table}
\egroup

For each sub-graph with an argument this argument corresponds to the target argument ($c_{tar}$) of the last node in the sub-graph. For example, let $E = \texttt{1+(2+3)}$ then the graph of $\subt{E}(t)$ is recursively constructed as in graph \ref{graph:exexpr}. Notice how the argument, $t$, of the sub-graph is the third argument of the last $\mathit{bop}_+(\_)$ node.

\begin{definition}
\label{def:cfg}
A control-flow graph is a four-tuple $G = (V, E, s, t)$, where $V$ is a set of nodes, $E$ is a set of node pairs representing an edge between two nodes, $s\in V$ is a start node and $t\in V$ is an exit node.
\end{definition}

When illustrated as a graph, the entry node of a control-flow graph is marked with an in-going edge with no origin and the exit node is marked with an outgoing edge with no target. In example graph \ref{graph:exexpr}, $s = \mathit{constRead}(1, t_1)$ and $t = \mathit{bop}_+(t_1, t_2, t)$.

There are seventeen different nodes, all introduced below
\begin{description}
\item{$\mathit{start}$:} This node indicates the start of a program and is the first node of any program or function body. 
\item{$\mathit{bop_{\oplus}}(t_l,t_r, t_{tar})$, $\mathit{sop_{\oplus}}(t_l,t_r, t_{tar})$, $\mathit{uop_{\circ}}(t, t_{tar})$, $\mathit{inc_{\circ}}(h, t_{tar})$:} These nodes indicate binary, short-circuit-binary, unary, and increment/decrement operations, respectively. The operations of the first three are all performed on temporary storage, while the fourth is performed on the heap. The last argument, $t_{tar}$, indicates where the result of performing the operation should be stored. For example $\texttt{1+2}$ is a binary operation ($\mathit{bop}$) returning $\texttt{3}$, this value should be stored in temporary storage at $t_{tar}$. The arguments $t_l$ and $t_r$ correspond to entries in the temporary storage containing the values of the left operand and the right operand, respectively. For unary operations the $t$ argument is where the value of the operand is stored in temporary storage and for the increment and decrement operations the $h$ variable is the set of heap-locations storing possible operand values.
\item{$\mathit{if}(t)$:} This node has one in-going and two out-going edges, representing the choice of one branch or the other. The argument $t$ is the entry in temporary storage where the evaluated value of the condition is stored.
\item{$\mathit{constRead}(k, t_{tar})$:} This node represents reading a constant into the temporary storage at $t_{tar}$. The constant $k$ can be a string, a boolean, null or a number.
\item{$\mathit{varRead}(\texttt{\$v}, c_{tar})$, $\mathit{varWrite}(\texttt{\$v}, c_{val}, t_{tar})$:} These nodes indicate reading from and writing to a variable \texttt{\$v} respectively. Depending on the context, the target of the read and value of the write can either be a temporary or temporary heap variable name. 
\item{$\mathit{arrayInit}(t_{tar})$:} This node represents initializing an empty array in the temporary storage at $t_{tar}$.
\item{$\mathit{arrayAppend}(t_{val}, t_{arr})$, $\mathit{arrayAppend}(h_{var}, c_{val}, t_{tar})$, $\mathit{arrayAppend}(h_{var}, h_{tar})$:} With three different signatures, this node represents appending to an array in temporary storage, in the heap, or in a read-context. Appending to an array in temporary storage happens when an array is initialized with values without keys, in which case the first type of append node is used. Appending to an array on the heap is an ordinary array append, where $h_{var}$ is the possible heap locations of the array, $c_{val}$ is either a heap temporary variable name in case of a reference append or a temporary variable name in case of a normal append. The value of $c_{val}$ is stored in $t_{tar}$. Appending in a read context, i.e.\ without a value to be written, is usually not allowed, however there are some exceptions. For those cases the last node is used and the appended heap location is placed in temporary heap storage for later access.
\item{$\mathit{arrayRead}(h_{arr}, t_{key}, c_{tar})$:} Reading from an array on the heap ($h_{arr}$) with a key derived from the value at $t_{key}$. The result can either be a reference to the entry or the value at the given key, hence the $c_{tar}$ variable. 
\item{$\mathit{arrayWrite}(t_{val}, t_{arr})$, $\mathit{arrayWrite}(h_{arr}, c_{val}, t_{tar})$:} Writing to an array in temporary storage or in the heap respectively. The arguments to the former ($t_{val}$ and $t_{arr}$) are temporary variable names for the value and the array respectively. Just as with append, the value written can be either a reference or a value, hence the $c_{val}$ variable. The value evaluated from $c_{val}$ is stored in temporary storage at $t_{tar}$.
\item{$\mathit{call}_{\mathit{fn}}(c_1, \cdots, c_n)$, $\mathit{exit}(c_1, \cdots, c_m)$, $\mathit{result}_{v}(c_{tar})$:} A function call is performed with the $\mathit{call}$ node. This holds the name, $\mathit{fn}$, of the function being called (which can always be resolved) as well as information about function parameters ($c_1, \cdots, c_n$). For every call node, $v$, there is a single result node, $\mathit{result}_v$. This restores the calling context and stores the result of the function call in $c_{tar}$. The node preceding the result node is an $\mathit{exit}$ node, which is unique to the function being called and the last node of the function sub-graph. The exit arguments $c_1, \cdots, c_m$ are entries in the temporary storage which contain the evaluations of return statements in the function.
\item{$nop$:} This node does nothing and is only there for structural purposes. It is in the control-flow graphs denoted as a small node with no label.
\end{description}

\begin{graph}
\centering
\begin{adjustbox}{max size={1\textwidth}{.25\textheight}, margin = 10pt}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
	\node [entry] (en) {};
 	\node [subtree, right of =en] (t1){\scriptsize $\subt{\texttt{1+(2+3)}}(t)$};
	\node [exit, right of=t1] (ex) {};
    % Draw edges
    \path[line] (en) -> (t1);
    \path[line] (t1) ->(ex);

\end{tikzpicture}\end{adjustbox}

\begin{adjustbox}{max size={1\textwidth}{.25\textheight}, margin = 10pt}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
	\node [entry] (en) {};
 	\node [node, right of=en, label={270:$\mathit{constRead}(\texttt{1},t_1)$}] (c1) {};
 	\node [subtree, right of =c1] (t1){$\subt{\texttt{2+3}}(t_2)$};
 	\node [node, right of=t1, label={270:$\mathit{bop}_+(t_1,t_2, t)$}] (bop1) {};
	\node [exit, right of=bop1] (ex) {};
    % Draw edges
    \path[line] (en) -> (c1);
    \path[line] (c1) -> (t1);
    \path[line] (t1) -> (bop1);
     \path[line] (bop1) -> (ex);

\end{tikzpicture}\end{adjustbox}
\begin{adjustbox}{max size={1\textwidth}{.25\textheight}, margin = 10pt}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
	\node [entry] (en) {};
 	\node [node, right of=en, label={270:$\mathit{constRead}(\texttt{1},t_1)$}] (c1) {};
 	\node [node, right of=c1, label={90:$\mathit{constRead}(\texttt{2},t_3)$}] (c2) {};
 	\node [node, right of=c2, label={270:$\mathit{constRead}(\texttt{3},t_4)$}] (c3) {};
 	\node [node, right of=c3, label={90:$\mathit{bop}_+(t_3,t_4, t_2)$}] (bop1) {};
 	\node [node, right of=bop1, label={270:$\mathit{bop}_+(t_1,t_2, t)$}] (bop2) {};
	\node [exit, right of=bop2] (ex) {};
    % Draw edges
    \path[line] (en) -> (c1);
    \path[line] (c1) -> (c2);
    \path[line] (c2) -> (c3);
    \path[line] (c3) -> (bop1);
     \path[line] (bop1) -> (bop2);
    \path[line] (bop2) -> (ex);

\end{tikzpicture}\end{adjustbox}

\caption{Recursively constructing graph $\subt{\texttt{1+(2+3)}}(t)$}
\label{graph:exexpr}
\end{graph}

After parsing a P0 program $p : \syn{\mathit{program}}$ to an abstract syntax tree the control-flow graph, $G_p = (V_p, E_p, s_p, t_p)$, can be constructed recursively by constructing sub-graphs for statements $S_1, \cdots, S_n$ in $p$. The sub graphs, $G_i = (V_i, E_i, s_i, t_i)$ and $G_{i+1} = (V_{i+1}, E_{i+1}, s_{i+1}, t_{i+1})$, for statement $i$ and $i+1$ respectively are connected by edges from $t_i$ to $s_{i+1}$. A separate $\mathit{start}$ node is created as $s_p$ and $t_p = t_n$. For each function definition in $p$ a separate graph, $G_f = (V_f, E_f, s_f, t_f)$, is created with $s_f = \mathit{start}_f$ and $t_f = \mathit{exit}_f$. Function calls in $p$ create edges to the corresponding $s_f$ and from the corresponding $t_f$.

\subsection{Statements}
Let $s : \syn{statement}$ be a statement, then there are nine different graphs, one for each case in the grammar \ref{gramm:p0}. The first four, for $s = $ \texttt{while($E$) $S$}, $s = $ \texttt{for($E_1$;$ E_2$;$E_3$) $S$}, $s = $ \texttt{if($E$) $S$}, and $s = $ \texttt{if($E$) $S_1$ else $S_2$} statements, are depicted as graph \ref{graph:flowwhile}, \ref{graph:flowfor}, \ref{graph:flowif}, and \ref{graph:flowifelse} respectively.


\begin{graph}
\captionsetup{justification=centering}
\hspace*{\fill}
\subcaptionbox{\texttt{while}-statement\label{graph:flowwhile}}{\begin{adjustbox}{max size={.25\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{E}(t)$};
    \node [node, below of=e1, label={0:$\mathit{if}(t)$}] (if) {};
    \node [subtree, below of=if ] (s1) {$\subt{S}$};
    \node [emptynode, below of=s1 ] (end) {};
    \node [exit, below of=end] (ex) {};

    % Draw edges
    \path[line] (en) -> (e1);
    \path [line] (e1) -> (if);
    \path [line] (if) -> (s1);
    \path [line] (if) -- (-2, -4) |- (end); 
    \path [line] (s1) -- (2, -6) |- (e1); 
    \path [line] (end) -> (ex);
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{\texttt{for}-statement\label{graph:flowfor}}{\begin{adjustbox}{max size={.25\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{E_1}(\_)$};
    \node [subtree, below of=e1] (e2) {$\subt{E_2}(t)$};
    \node [node, below of=e2, label={135:$\mathit{if}(t)$}] (if) {};
    \node [subtree, below left of=if] (s) {$\subt{S}$};
    \node [subtree, below right of=if] (e3) {$\subt{E_3}(\_)$};
    \node [emptynode, below right of=s] (end) {};
    \node [exit, below of=end] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (e1);
    \path [line] (e1) -> (e2);
    \path [line] (e2) -> (if);
    \path [line] (if) -> (s);
    \path [line] (s) -> (e3);
%    \path [line] (e3) -> (end);
    \path [line] (end) -> (ex);
    \path [line] (e3) |- (e2);
    \path [line] (if) -- (-3, -6) |- (end);
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{\texttt{if}-statement\label{graph:flowif}}{\begin{adjustbox}{max size={.25\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{E}(t)$};
    \node [node, below of=e1, label={45:$\mathit{if}(t)$}] (if) {};
    \node [subtree, below of=if] (s1) {$\subt{S}$};
    \node [emptynode, below of=s1](end) {};
    \node [exit, below of=end] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (e1);
    \path [line] (e1) -> (if);
    \path [line] (if) -> (s1);
    \path [line] (s1) -> (end);
    \path [line] (if) -- (2,-4) |-  (end);
    \path [line] (end) -> (ex);
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{\texttt{if-else}-statement\label{graph:flowifelse}}{\begin{adjustbox}{max size={.25\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree,below of=en] (e1) {$\subt{E}(t)$};
    \node [node, below of=e1, label={35:$\mathit{if}(t)$}] (if) {};
    \node [subtree, below left of=if] (s1) {$\subt{S_1}$};
    \node [subtree, below right of=if] (s2) {$\subt{S_2}$};
    \node [emptynode, below right of=s1](end) {};
    \node [exit, below of=end] (ex) {};
    % Draw edges
    \path[line] (en) -> (e1);
    \path [line] (e1) -> (if);
    \path [line] (if) -> (s1);
    \path [line] (if) -> (s2);
    \path [line] (s1) -> (end);
    \path [line] (s2) -> (end);
    \path [line] (end) -> (ex);
\end{tikzpicture}\end{adjustbox}}
\hspace*{\fill}
\caption{Statement graphs}
\end{graph}

%For $s =$ \texttt{global $v_1,\cdots, v_n$} the graph is containing a single, $global$, node which takes the variable names $v_1,\cdots v_n$ as argument. 
Return-statements, $s =$ \texttt{return $T$}, have three different graphs. An empty statement, i.e.\ which does not contain any expressions, is equivalent to returning \texttt{null}. This case yields the graph in \ref{graph:floweret}. If $T : \syn{rexpr}$ and the function is a reference function, i.e.\ the function signature has an ampersand preceding the function name (see program \ref{lst:retref}), then the references are returned (graph \ref{graph:flowret} where $c : \text{HeapTemps}$). If this is not the case, (see program \ref{lst:retnoref}), and assuming the program is a valid P0 program, then it is fair to assume that $T$ can be parsed as an expression, since it cannot be an array-append operation. Assuming it was an array append operation, the program would not be a valid PHP program and thus not a valid P0 program (see section \ref{sec:langsubset}). Assuming $T : \syn{expr}$ the graphs for a return statement evaluates $T$ and return the result. This is the case in graph \ref{graph:flowret} with $c : \text{Temps}$.  For all graphs, the exit-node is the unique exit-node of the function.

\begin{program}
 \begin{subfigure}[b]{.45\linewidth}
	\begin{lstlisting} 
function normRet(){
	$a++;	
	return $a;
}
	\end{lstlisting}
    \caption{Normal return-statement}
    \label{lst:retnoref}
  \end{subfigure}\hfill%
 \begin{subfigure}[b]{.45\linewidth}
	\begin{lstlisting} 
function &refRet(&$a){
	return $a[];
}

	\end{lstlisting}
    \caption{Reference return-statement}
    \label{lst:retref}
  \end{subfigure}

\caption{Return-statement examples}
\label{lst:return-statements}
\end{program}

\begin{graph}
\captionsetup{justification=centering}
\hspace*{\fill}
\subcaptionbox{\texttt{return}-statement\label{graph:floweret}}{\begin{adjustbox}{max size={.3\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [node, below of=en, label={45:$\mathit{constRead}(\texttt{null}, t_{tar})$}] (const) {};
    \node [node, below of=const, label={45:$\mathit{exit}(\cdots, t_{tar}, \cdots)$}] (exitn) {};
    \node [emptynode, below of=exitn ] (end) {};
    \node [exit, below of=end] (ex) {};
      % Draw edges
    \path[line] (en) -> (const);
    \path[line] (const) -> (exitn);
    \path[line] (end) -> (ex);
    
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{\texttt{return}-expression-statement\label{graph:flowret}}{\begin{adjustbox}{max size={.3\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (exp) {$\subt{T}(c)$};
    \node [node, below of=exp, label={45:$\mathit{exit}(\cdots, c, \cdots)$}] (exitn) {};
    \node [emptynode, below of=exitn ] (end) {};
    \node [exit, below of=end] (ex) {};
    % Draw edges
    \path[line] (en) -> (exp);
    \path[line] (exp) -> (exitn);
    \path[line] (end) -> (ex);
    
\end{tikzpicture}\end{adjustbox}}
\hspace*{\fill}
\caption{Return-statement graphs}
\end{graph}

The remaining four graphs are the empty graph, the graph of the expression statement, the graph of \texttt{global} statement, and the graph of the block statement, which are all straight-forward.


\subsection{Expressions}

Let $e : \syn{expr}$, then by ignoring the trivial parenthesized expression case, viewing the four increment/decrement operations as one, and the two array-initialization operations as one, the expression can be of nine different forms. For $e = E_l \oplus E_r$, the graph depends on the operation. If the operation is a short-circuit operation, logical \texttt{\&\&} or \texttt{||}, then the graph is as \ref{graph:flowsop}, since only one branch may be required to be evaluated. If the operation is not a short-circuit operation, then the graph becomes as \ref{graph:flowbop}, since both expressions must be evaluated. In both cases, the operations are performed on and saved in the temporary storage. The unary operation $e = \circ E$ is similar to the previous case with a graph as \ref{graph:flowuop}. A separate graph for unary post/pre increment/decrement operations is necessary, because of the performed update on the heap location. This is the reason for the operations not being performed on the temporary storage, but instead on heap-locations directly. The result of the operation is stored in the temporary storage. Figure \ref{graph:flowinc} illustrates the corresponding flow graph.
\begin{graph}
\captionsetup{justification=centering}
\hspace*{\fill}
\subcaptionbox{Short-circuit binary operation\label{graph:flowsop}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{E_l}(t_l)$};
    \node [node, below of=e1, label={0:$\mathit{if}(t_l)$}] (if) {};
    \node [subtree, below of=if] (e2) {$\subt{E_r}(t_r)$};
    \node [node, below of=e2, label={0:$\mathit{sop}_\oplus(t_l,t_r,t_{tar})$}] (sop) {};
    \node [exit, below of=sop] (ex) {};

    % Draw edges
    \path [line] (en) -> (e1);
    \path [line] (e1) -> (if);
    \path [line] (if) -> (e2);
    \path [line] (e2) -> (sop);
    \path [line] (if) -- (-2, -4) |- (sop);
    \path [line] (sop) -> (ex);
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Binary operation\label{graph:flowbop}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{E_l}(t_l)$};
    \node [subtree, below of=e1] (e2) {$\subt{E_r}(t_r)$};
    \node [node, below of=e2, label={0:$\mathit{bop}_\oplus(t_l,t_r,t_{tar})$}] (bop) {};
    \node [exit, below of=bop] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (e1);
    \path [line] (e1) -> (e2);
    \path [line] (e2) -> (bop);
    \path [line] (bop) -> (ex);
    
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Unary operation\label{graph:flowuop}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e) {$\subt{E}(t_{val})$};
    \node [node, below of=e, label={0:$\mathit{uop}_\circ(t_{val},t_{tar})$}] (uop) {};
    \node [exit, below of=uop] (ex) {};

    % Draw edges
    \path [line] (en) -> (e);
    \path [line] (e) -> (uop);
    \path [line] (uop) -> (ex);
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Post/pre increment/decrement opreation\label{graph:flowinc}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e) {$\subt{V}(h_{val})$};
    \node [node, below of=e, label={0:$\mathit{inc}_\circ(h_{val},t_{tar})$}] (inc) {};
    \node [exit, below of=inc] (ex) {};

    % Draw edges
    \path [line] (en) -> (e);
    \path [line] (e) -> (inc);
    \path [line] (inc) -> (ex);
\end{tikzpicture}\end{adjustbox}}%
\hspace*{\fill}
\caption{Operation-expression graphs}
\end{graph}

For a variable read $e = \texttt{\$v}$ (for some variable \texttt{\$v}), or a constant read, e.g.\ $e = \texttt{"foo"}$, the graph is a single $\mathit{varRead}(\texttt{\$v}, t_{tar})$ or $\mathit{constRead}(\texttt{"foo"}, t_{tar})$ node, respectively. For an array read expression, $e = E_{arr}\texttt[E_{key}\texttt{]}$, the sub array expression, $E_{arr}$, should be evaluated before the key expression, $E_{key}$, and the graph then becomes like graph \ref{graph:flowaread}, where $T$ is the graph corresponding to $E_{arr}$ and $c_{arr}, c_{tar} : \text{Temps}$. If the expression is an array initialization, $e = \texttt{[}\cdots, E_1, \cdots, E_2\texttt{=>}E_3\texttt{]}$, then an array is first initialized in temporary storage after the entries are either appended or written to the array, hence graph \ref{graph:flowai}.
\begin{graph}
\captionsetup{justification=centering}
\hspace*{\fill}
\subcaptionbox{Array read \label{graph:flowaread}}{\begin{adjustbox}{max size={.3\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{T}(c_{arr})$};
    \node [subtree, below of=e1] (e2) {$\subt{E_{key}}(t_{key})$};
    \node [node, below of=e2, label={0:$\mathit{arrayRead}(c_{ar},t_{key},c_{tar})$}] (rd) {};
    \node [exit, below of=rd] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (e1);
    \path [line] (e1) -> (e2);
    \path [line] (e2) -> (rd);
    \path [line] (rd) -> (ex);
\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Array initialize\label{graph:flowai}}{\begin{adjustbox}{max size={.3\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [node, below of=en, label={0:$\mathit{arrayInit}(t_{arr})$}] (ai) {};
    \node [dots, left of=ai] (dots1) {$\vdots$};
    \node [subtree, below of=dots1] (e1) {$\subt{E_1}(t_1)$};
    \node [node, right of=e1, label={45:$\mathit{arrayAppend}(t_1, t_{arr})$}] (app) {};
	\node [dots, right of=app] (dots2) {$\vdots$};
    \node [subtree, below of=dots2] (e2) {$\subt{E_2}(t_2)$};
    \node [subtree, left of=e2] (e3) {$\subt{E_3}(t_2)$};
	\node [node, left of=e3, label={-85:$\mathit{arrayWrite}(t_2, t_3, t_{arr})$}] (aw) {};
	\node [exit, below of=aw] (ex) {};

    % Draw edges
    \path [line] (en) -> (ai);
    \path [line] (ai) -> (dots1);
    \path [line] (dots1) -> (e1);
    \path [line] (e1) -> (app);
    \path[line] (app) -> (dots2);
    \path[line] (dots2) -> (e2);
    \path[line] (e2) -> (e3);
    \path[line] (e3) -> (aw);
        \path[line] (aw) -> (ex);



\end{tikzpicture}\end{adjustbox}}
\hspace*{\fill}
\caption{Array-read and -initialize graphs}
\end{graph}

For function calls $e = \mathit{fn}\texttt{(}T_1,\cdots, T_n\texttt{)}$, the result variable is a temporary variable, $c_{tar}:\text{Temps}$, the arguments are either an expression, $T_i : \syn{expr}$ and $c_i : \text{Temps}$, or a reference expression, $T_i : \syn{rexpr}$ and $c_i : \text{HeapTemps}$, depending on the signature of $\mathit{fn}$. If the argument is pass-by-reference, i.e.\ the variable $v_i$ is denoted with a \texttt{\&} in the signature, then the expression must be a reference expression. If not, the argument is an expression. This follows from the program being a valid P0 program. The function graph will be as graph \ref{graph:flowcall}, where the start node, exit node, and function body are the unique nodes of the $\mathit{fn}$ function, i.e.\ they are not copied. This ensures that the graph is finite, but multiple, say $n$, calls to the same function will yield a graph with $n$ edges to the start node and from the exit node. Notice the dashed line going directly from the call node to the return node. This indicates that the call node may pass information directly to the return node, used for restoring the local context before calling $\mathit{fn}$.

\begin{graph}
\centering
\captionsetup{justification=centering}
\begin{adjustbox}{max size={.3\textwidth}{.40\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e1) {$\subt{T_1}(c_1)$};
    \node [dots, below of=e1] (dots) {$\vdots$};
    \node [subtree, below of=dots] (em) {$\subt{T_n}(c_n)$};
    \node [node, below of=em] [label={-90:$call_{\mathit{fn}}(c_1,\cdots,c_n)$}] (call) {};
    \node [node, right of=call, label={$\mathit{start}$}] (entry) {};
    \node [subtree, below of=entry] (sl) {$B$};
    \node [node, below of=sl, label={-90:$\mathit{exit}(t)$}] (exit) {};
    \node [node, left of=exit, label={$\mathit{result}_{call_{\mathit{fn}}}(c_{tar})$}] (result) {};
    \node [exit, below of=result] (ex) {};

    % Draw edges
    \path [line] (en) -> (e1);
    \path [line] (e1) -> (dots);
    \path [line] (dots) -> (em);
    \path [line] (em) -> (call);
    \path [line] (call) -> (entry);
    \path [line] (entry) -> (sl);
    \path [line] (sl) -> (exit);
    \path [line] (exit) -> (result);
    \path [line] (result) -> (ex);
    \draw [dashed, ->] (call) -- (-2, -8) |- (result);
\end{tikzpicture}\end{adjustbox}
\caption{Function-call graph \label{graph:flowcall}}
\end{graph}

Finally, the expression can be an assignment. There are two types of assignments, a regular value-assignment and a reference-assignment, using the \texttt{\&} operator. Each assignment can be split into three categories; variable-, array-write-, and array-append-assignments, depending on the target (left-hand side of the operation). Examples of these six operations can be seen in program \ref{lst:assignments}. Due to the distinction between temporary storage of values and heap-location-sets, these six cases must be handled individually. For value array write ($e = R\texttt{[$E_{key}$]=} E_{val}$), array append ($e = R\texttt{[]=} E_{val}$), and variable write ($e = \texttt{\$v=} E_{val}$), the graphs are as shown in graphs \ref{graph:flowawrite}, \ref{graph:flowaappend}, and \ref{fig:vwrite}, respectively, where $T$ is the subgraph of the value expression $E_{val} : \syn{expr}$ and $c_{val} : \text{Temps}$ is the temporary variable holding the result. The reference assignments $e = R\texttt{[$e_{key}$]=\&} R_{val}$, $e = r\texttt{[]=\&} R_{val}$, and $e = \texttt{\$v=\&} R_{val}$ are similar, but with $T$ being the subgraph of the reference expression, $R_{val} : \syn{rexpr}$ and $c_{val} : \text{HeapTemp}$ the heap temporary variable holding the heap locations of the value.
\begin{graph}
\captionsetup{justification=centering}
\hspace*{\fill}
\subcaptionbox{Array write\label{graph:flowawrite}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (ve) {$\subt{R}(h_{var})$};
    \node [subtree, below of=ve] (e1) {$\subt{E}(t_{key})$};
    \node [subtree, below of=e1] (e2) {$\subt{T}(c_{val})$};
    \node [node, below of=e2, label={0:$\mathit{arrayWrite}(h_{var}, t_{key},c_{val},t_{tar})$}] (rd) {};
    \node [exit, below of=rd] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (ve);
    \path [line] (ve) -> (e1);
    \path [line] (e1) -> (e2);
    \path [line] (e2) -> (rd);
    \path [line] (rd) -> (ex);

\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Array append \label{graph:flowaappend}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (ve) {$\subt{R}(h_{var})$};
    \node [subtree, below of=ve] (e1) {$\subt{T}(c_{val})$};
    \node [node, below of=e1, label={0:$\mathit{arrayAppend}(h_{var}, c_{val},t_{tar})$}] (rd) {};
    \node [exit, below of=rd] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (ve);
    \path [line] (ve) -> (e1);
    \path [line] (e1) -> (rd);
    \path [line] (rd) -> (ex);

\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Array append without write\label{graph:flowaappend2}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (ve) {$\subt{R}(h_{var})$};
    \node [node, below of=ve, label={0:$\mathit{arrayAppend}(h_{var},h_{tar})$}] (rd) {};
    \node [exit, below of=rd] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (ve);
    \path [line] (ve) -> (rd);
    \path [line] (rd) -> (ex);

\end{tikzpicture}\end{adjustbox}}\hfill%
\subcaptionbox{Variable write\label{fig:vwrite}}{\begin{adjustbox}{max size={.24\textwidth}{.25\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [entry] (en) {};
    \node [subtree, below of=en] (e2) {$\subt{T}(c_{val})$};
    \node [node, below of=e2, label={0:$\mathit{varWrite}(\texttt{\$v},c_{val},t_{tar})$}] (ass) {};
    \node [exit, below of=ass] (ex) {};
    
    % Draw edges
    \path [line] (en) -> (e2);
    \path [line] (e2) -> (ass);
    \path [line] (ass) -> (ex);
\end{tikzpicture}\end{adjustbox}}
\hspace*{\fill}
\caption{Array-append, array-write and variable-write expressions}
\end{graph}
\begin{program}
\begin{lstlisting}
<?php

$a = []; //Value assignment
$a[] = 1; //Array append assignment
$a[1] = 2; //Array write assignment

$b = &$a; //Value reference assignment
$c[] = &$a;//Array append reference assignment
$d[1] = &$a; //Array write reference assignment

\end{lstlisting}
\caption{Assignments}
\label{lst:assignments}
\end{program}

\subsection{Reference and variable expressions}
Since PHP supports nested assignments, e.g.\ \texttt{\$a[]['foo']  = \&func()[]}, deciding which location to update is performed recursively. The graph of this assignment is illustrated as graph \ref{graph:exassign}. Here the locations of variable \texttt{\$a} is stored in variable $h_{var'}$. A new location $l$ is then appended to the array values of the locations in $h_{var'}$, and $h_{var} = \{l\}$. Now the expression of the index is resolved and stored in $t_{key}$, and the locations of the right-hand side are resolved and stored in $h_{val}$. Finally, the locations in $h_{val}$ are written to the array at $h_{var}$, i.e.\ $l$, with the key of $t_{key}$. 

\begin{graph}
\centering
\begin{adjustbox}{max size={.33\textwidth}{.33\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
	\node [entry] (en) {};
	\node [subtree, below of =en](t1) {	 $\subt{\texttt{\$$$a[]}}(h_{var})$};
	\node [subtree, below of =t1](t2) {$\subt{\texttt{'foo'}}(t_{key})$};
	\node [subtree, below of =t2](t3) { $\subt{\texttt{func()[]}}(h_{val})$};
	\node [node, below of =t3, label={-45:$\mathit{arrayWrite}(h_{var}, t_{key}, h_{val}, t)$}] (ass) {};
	\node [exit, below of=ass] (ex) {};
    % Draw edges
    \path [line] (en) -> (t1);
    \path [line] (t1) -> (t2);
    \path [line] (t2) -> (t3);
    \path [line] (t3) -> (ass);
    \path [line] (ass) -> (ex);

\end{tikzpicture}\end{adjustbox}
\begin{adjustbox}{max size={.33\textwidth}{.33\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
	\node [entry] (en) {};
	\node [subtree, below of =en](t1) {	 $\subt{\texttt{\$$$a}}(h_{var'})$};
	\node [node, below of = t1, label={-45:$\mathit{arrayAppend}(h_{var'}, h_{var})$}] (aa) {};
	\node [node, below of = aa, label={-45:$\mathit{constRead}(\texttt{'foo'}, t_{key})$}] (cr) {};		
	\node [subtree, below of =cr](t3) { $\subt{\texttt{func()}}(h_{val'})$};
	\node [node, below of = t3, label={-45:$\mathit{arrayAppend}(h_{val'}, h_{val})$}] (aa2) {};
	\node [node, below of =aa2, label={-45:$\mathit{arrayWrite}(h_{var}, t_{key}, h_{val}, t)$}] (ass) {};
	\node [exit, below of=ass] (ex) {};
    % Draw edges
    \path [line] (en) -> (t1);
    \path [line] (t1) -> (aa);
    \path [line] (aa) -> (cr);
    \path [line] (cr) -> (t3);
    \path [line] (t3) -> (aa2);
    \path [line] (aa2) -> (ass);
    \path [line] (ass) -> (ex);

\end{tikzpicture}\end{adjustbox}
\begin{adjustbox}{max size={.33\textwidth}{.33\textheight}}\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
	\node [entry] (en) {};
	\node [node, below of =en, label = {-45: $\mathit{varRead}(\texttt{\$$$a}, h_{var'})$}](t1) {};
	\node [node, below of = t1, label={-45:$\mathit{arrayAppend}(h_{var'}, h_{var})$}] (aa) {};
	\node [node, below of = aa, label={-45:$\mathit{constRead}(\texttt{'foo'}, t_{key})$}] (cr) {};		
	\node [node, below of =cr, label={-45:$call_{\texttt{func}}()$}](call) {};
	\node [node, below of =call, label={-45:$result_{call_{\texttt{func}}()}(h_{val'})$}](t3) {};
	\node [entry, right of =call] (fn) {};
	\node [node, below of = t3, label={-45:$\mathit{arrayAppend}(h_{val'}, h_{val})$}] (aa2) {};
	\node [entry, right of =t3] (fex) {};
	\node [node, below of =aa2, label={-45:$\mathit{arrayWrite}(h_{var}, t_{key}, h_{val}, t)$}] (ass) {};
	\node [exit, below of=ass] (ex) {};
    % Draw edges
    \path [line] (en) -> (t1);
    \path [line] (t1) -> (aa);
    \path [line] (aa) -> (cr);
    \path [line] (cr) -> (call);
    \path [line] (call) -> (fn);
    \draw [dashed, ->] (call) -- (t3);
    \path [line] (t3) -> (aa2);
    \path [line] (fex) -> (t3);
    \path [line] (aa2) -> (ass);
    \path [line] (ass) -> (ex);

\end{tikzpicture}\end{adjustbox}
\caption{Creating graph of expression $\subt{\texttt{\$a[]['foo']  = \&func()[]}}(t)$}
\label{graph:exassign}
\end{graph}

The sub-graphs which take a $h : \text{HeapTemp}$ variable are called reference- and variable-expressions, where the only difference between the two is that reference expressions may contain function references. As seen in the previous example, they resolve heap-location sets rather than values. For $r = \mathit{fn} \texttt{(}c_1, \cdots, c_n\texttt{)}$ the graph corresponds to graph \ref{graph:flowcall}, with a heap-location set as the result, i.e.\ $c_{tar} : \text{HeapTemps}$. For the variable read $r = \texttt{\$v}$ the graph is a single $\mathit{varRead}(\texttt{\$v}, h_{tar})$ node. When the expression is an array-read $r= R_{arr}[E_{key}]$ the graph corresponds to that of graph \ref{graph:flowaread} with $c_{arr},c_{tar} : \text{HeapTemps}$ and $T$ being the graph of $R_{arr}$. Finally, for the append operation $r = R_{arr}[]$ the graph corresponds to that of graph \ref{graph:flowaappend2}.